---
title: "Compute Emissions Dashboard"
freeze: false
---

## Purpose

This page summarizes tracked compute emissions for epi-ML runs and modules.

```{r}
#| echo: false
read_csv_safe <- function(path) {
  if (!file.exists(path)) return(data.frame())
  tryCatch(read.csv(path, stringsAsFactors = FALSE), error = function(e) data.frame())
}

events_path <- file.path("data", "public", "outputs", "emissions", "emissions_module_events.csv")
runs_path <- file.path("data", "public", "outputs", "emissions", "emissions_run_summary.csv")
modules_path <- file.path("data", "public", "outputs", "emissions", "emissions_module_summary.csv")
reco_path <- file.path("data", "public", "outputs", "emissions", "emissions_recommendations.csv")

events <- read_csv_safe(events_path)
runs <- read_csv_safe(runs_path)
modules <- read_csv_safe(modules_path)
recommendations <- read_csv_safe(reco_path)

has_data <- nrow(events) > 0 && nrow(modules) > 0 && nrow(runs) > 0
if (has_data) {
  events$timestamp_utc <- as.POSIXct(events$timestamp_utc, format = "%Y-%m-%dT%H:%M:%SZ", tz = "UTC")
  runs$start_time_utc <- as.POSIXct(runs$start_time_utc, format = "%Y-%m-%dT%H:%M:%SZ", tz = "UTC")
  runs$end_time_utc <- as.POSIXct(runs$end_time_utc, format = "%Y-%m-%dT%H:%M:%SZ", tz = "UTC")
  events <- events[order(events$run_id, events$elapsed_minutes_run), , drop = FALSE]
  modules <- modules[order(modules$total_emissions_kg, decreasing = TRUE), , drop = FALSE]
}
```

```{r}
#| echo: false
#| results: asis
if (!has_data) {
  cat(
    "No emissions data is available yet.\n\n",
    "Run `Rscript scripts/run_modules.R --track-emissions` locally, then render the site again.",
    sep = ""
  )
} else {
  total_emissions_kg <- sum(events$emissions_kg, na.rm = TRUE)
  total_energy_kwh <- sum(events$energy_kwh, na.rm = TRUE)
  total_hours <- sum(events$duration_seconds, na.rm = TRUE) / 3600
  total_runs <- length(unique(events$run_id))
  total_modules <- length(unique(events$module))
  total_exec <- nrow(events)
  mean_module_emissions_g <- mean(events$emissions_kg, na.rm = TRUE) * 1000

  cat(sprintf("<strong>Tracked runs:</strong> %d<br/>", total_runs))
  cat(sprintf("<strong>Total modules:</strong> %d<br/>", total_modules))
  cat(sprintf("<strong>Module executions:</strong> %d<br/>", total_exec))
  cat(sprintf("<strong>Total emissions:</strong> %.3f g CO2e<br/>", total_emissions_kg * 1000))
  cat(sprintf("<strong>Total energy:</strong> %.6f kWh<br/>", total_energy_kwh))
  cat(sprintf("<strong>Tracked compute time:</strong> %.2f minutes<br/>", total_hours * 60))
  cat(sprintf("<strong>Average per module execution:</strong> %.4f g CO2e<br/><br/>", mean_module_emissions_g))
}
```

For native dashboard UI (similiar to CodeCarbon `dashboard.png`), run:

```bash
carbonboard --filepath=logs/emissions/codecarbon_emissions.csv
```

CodeCarbon tracking during local module computations:

```bash
python3 -m pip install codecarbon
Rscript scripts/run_modules.R --track-emissions
```

CodeCarbon dashboard (optional) upload:

```bash
CODECARBON_SAVE_TO_API=1 CODECARBON_EXPERIMENT_ID=<experiment-id> Rscript scripts/run_modules.R --track-emissions
```

## Computational Runs Timeline

```{r}
#| echo: false
#| tbl-colwidths: [20,20,14,14,16,16]
if (has_data) {
  runs_ord <- runs[order(runs$start_time_utc), , drop = FALSE]
  offsets <- cumsum(c(0, head(runs_ord$total_duration_seconds / 60, -1)))
  names(offsets) <- runs_ord$run_id
  events_seq <- events
  events_seq$elapsed_minutes_global <- events_seq$elapsed_minutes_run + offsets[events_seq$run_id]
  events_seq <- events_seq[order(events_seq$elapsed_minutes_global), , drop = FALSE]
  events_seq$cumulative_emissions_g_global <- cumsum(events_seq$emissions_kg) * 1000
  x_max <- max(events_seq$elapsed_minutes_global, na.rm = TRUE)
  y_max <- max(events_seq$cumulative_emissions_g_global, na.rm = TRUE)
  plot(
    events_seq$elapsed_minutes_global,
    events_seq$cumulative_emissions_g_global,
    type = "l",
    lwd = 2,
    col = "#2F6B8A",
    xlim = c(0, x_max * 1.05),
    ylim = c(0, y_max * 1.05),
    xlab = "Elapsed minutes across runs (sequential)",
    ylab = "Cumulative emissions (g CO2e)",
    main = "Cumulative emissions timeline"
  )
  points(
    events_seq$elapsed_minutes_global,
    events_seq$cumulative_emissions_g_global,
    pch = 16,
    cex = 0.8,
    col = "#2F6B8A"
  )
  if (nrow(runs_ord) > 1) {
    run_boundaries <- offsets[-1]
    for (b in run_boundaries) {
      abline(v = b, lty = 3, col = "#9AA6B2")
    }
  }
} else {
  plot.new()
  title("No run timeline available")
}
```

## Average Module Duration Trend

```{r}
#| echo: false
if (has_data && nrow(events) > 0) {
  runs_ord <- runs[order(runs$start_time_utc), , drop = FALSE]
  offsets <- cumsum(c(0, head(runs_ord$total_duration_seconds / 60, -1)))
  names(offsets) <- runs_ord$run_id

  events_seq <- events
  events_seq$elapsed_minutes_global <- events_seq$elapsed_minutes_run + offsets[events_seq$run_id]
  events_seq$duration_minutes <- events_seq$duration_seconds / 60
  events_seq <- events_seq[order(events_seq$elapsed_minutes_global), , drop = FALSE]
  events_seq$cumulative_avg_duration_minutes <- cumsum(events_seq$duration_minutes) / seq_len(nrow(events_seq))

  overall_avg_duration <- mean(events_seq$duration_minutes, na.rm = TRUE)
  x_max <- max(events_seq$elapsed_minutes_global, na.rm = TRUE)
  y_max <- max(c(events_seq$cumulative_avg_duration_minutes, overall_avg_duration), na.rm = TRUE)
  upper_y <- if (is.finite(y_max) && y_max > 0) y_max * 1.10 else 1

  plot(
    events_seq$elapsed_minutes_global,
    events_seq$cumulative_avg_duration_minutes,
    type = "l",
    lwd = 2,
    col = "#2F6B8A",
    xlim = c(0, x_max * 1.05),
    ylim = c(0, upper_y),
    xlab = "Elapsed minutes across runs (sequential)",
    ylab = "Average duration (minutes)",
    main = "Average module duration"
  )
  abline(h = overall_avg_duration, lty = 2, lwd = 2, col = "#9AA6B2")
  legend(
    "bottomright",
    legend = c("Cumulative average duration", "Overall average duration"),
    col = c("#2F6B8A", "#9AA6B2"),
    lty = c(1, 2),
    lwd = c(2, 2),
    bty = "n"
  )
} else {
  plot.new()
  title("No average-duration profile available")
}
```

## Run Summary

```{r}
#| echo: false
#| tbl-colwidths: [24,24,12,12,14,14]
if (has_data) {
  view_runs <- runs[, c(
    "run_id", "start_time_utc", "end_time_utc",
    "total_duration_seconds", "total_energy_kwh", "total_emissions_kg",
    "avg_emissions_rate_kg_per_hour"
  )]
  view_runs$start_time_utc <- format(view_runs$start_time_utc, "%Y-%m-%d %H:%M:%S UTC")
  view_runs$end_time_utc <- format(view_runs$end_time_utc, "%Y-%m-%d %H:%M:%S UTC")
  view_runs$total_duration_seconds <- round(view_runs$total_duration_seconds, 2)
  view_runs$total_energy_kwh <- round(view_runs$total_energy_kwh, 6)
  view_runs$total_emissions_g <- round(view_runs$total_emissions_kg * 1000, 4)
  view_runs$avg_emissions_rate_g_per_hour <- round(view_runs$avg_emissions_rate_kg_per_hour * 1000, 4)
  view_runs <- view_runs[, c(
    "run_id", "start_time_utc",
    "total_duration_seconds", "total_energy_kwh",
    "total_emissions_g", "avg_emissions_rate_g_per_hour"
  )]
  names(view_runs) <- c(
    "Run ID", "Start",
    "s", "kWh",
    "g CO2e", "g CO2e/hr"
  )
  knitr::kable(
    view_runs,
    format = "pipe",
    align = rep("c", ncol(view_runs)),
    escape = FALSE
  )
}
```

Notation:

- `s`: total run duration in seconds.
- `kWh`: total energy consumed in kilowatt-hours.
- `g CO2e`: total emissions in grams of CO2-equivalent.
- `g CO2e/hr`: avg emissions rate for run (g CO2e per hour).

## Modules Summary

```{r}
#| echo: false
#| tbl-colwidths: [20,8,8,8,8,8,8,8,8,16]
if (has_data) {
  view <- modules[, c(
    "module_label", "page_href",
    "avg_duration_seconds", "total_duration_seconds",
    "mean_emissions_kg", "median_emissions_kg", "total_emissions_kg",
    "avg_energy_kwh", "total_energy_kwh",
    "emissions_share_pct", "emissions_rate_kg_per_hour"
  )]
  view$Module <- sprintf("<a href='%s'>%s</a>", view$page_href, view$module_label)
  view$`m̄` <- round(view$avg_duration_seconds / 60, 2)
  view$`Σm` <- round(view$total_duration_seconds / 60, 2)
  view$`ḡ` <- round(view$mean_emissions_kg * 1000, 4)
  view$`g̃` <- round(view$median_emissions_kg * 1000, 4)
  view$`Σg` <- round(view$total_emissions_kg * 1000, 4)
  view$`W̄h` <- round(view$avg_energy_kwh * 1000, 3)
  view$`ΣWh` <- round(view$total_energy_kwh * 1000, 3)
  view$`%` <- round(view$emissions_share_pct, 2)
  view$`g CO2e/hr` <- round(view$emissions_rate_kg_per_hour * 1000, 4)
  view_out <- view[, c(
    "Module",
    "m̄", "Σm",
    "ḡ", "g̃", "Σg",
    "W̄h", "ΣWh",
    "%", "g CO2e/hr"
  )]
  knitr::kable(
    view_out,
    format = "pipe",
    align = rep("c", ncol(view_out)),
    escape = FALSE
  )
}
```

Notation:

- `m̄`: average duration per module execution (minutes).
- `Σm`: cumulative duration across executions (minutes).
- `ḡ`: arithmetic mean emissions per execution (grams CO2e).
- `g̃`: median emissions per execution (grams CO2e).
- `Σg`: cumulative emissions across executions (grams CO2e).
- `W̄h`: average energy per execution (watt-hours).
- `ΣWh`: cumulative energy across executions (watt-hours).
- `%`: module share of total tracked emissions (percentage).
- `g CO2e/hr`: emissions intensity in grams CO2e per hour.

## Reduction Recommendations

```{r}
#| echo: false
if (has_data && nrow(recommendations) > 0) {
  names(recommendations) <- c("Priority", "Recommendation", "Rationale")
  knitr::kable(recommendations)
}
```
