---
title: "Module: Power Design"
---

```{r}
#| echo: false
#| warning: false
#| message: false
source("_code_helpers.R")
```

## Script Metadata

```{r}
#| echo: false
#| results: asis
render_script_metadata("surveillance/investigation/05_power_design.R")
```

## Declared Outputs and Current Status

```{r}
#| echo: false
#| results: asis
render_outputs_table("surveillance/investigation/05_power_design.R")
```

## Power Focus

Power planning in EPI-ML is centered on survey-weighted `cannabis_any_use` (alias `cua`) × gender (self-reported) interaction targets from CPADS, with primary endpoints tied to alcohol outcomes (`heavy_drinking_30d` alias `hd30d`, `ebac_legal`, and continuous `ebac_tot`). The primary design target is Woman-versus-Man interaction power with near-equal subgroup allocation, while Transgender/Non-binary strata are retained explicitly as feasibility flags rather than used to set the primary trial-planning sample size. CPADS PUMF does not provide complete sex-at-birth fields in this workflow, so subgroup planning is anchored to gender (self-reported). Downstream modules are power-gated from these targets before observational model execution.

## Source Code

```{r}
#| echo: false
#| results: asis
render_script_code("surveillance/investigation/05_power_design.R")
```
